Comparative Evaluation of Spark and Flink Stream Processing

### 2017-03-10 paper/paper.tex
[paper] spell checking

\end{abstract}
Recent years have witnessed advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For instance, both frameworks support real-time data 
stream processing. The question of which platform is superior to the other still open. 
In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink by measuring different performance metrics (latency and throughput). Furthermore, we cover some key aspects of streaming data applications and how they are handled in the two frameworks. The experiment is done using multiple streaming workloads over real-world datasets. Results show that Flink outperform Spark streaming in term of latency. While Spark Streaming gives higher throughput rates than Flink. 
\section{Introduction}

\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia. This attention is a result of the fact that the amount of data we generate is increasing exponentially \cite{idc}.  Different sources such as  social media platforms, mobile phones, Internet of things (IoT), financial transactions, etc.  create large volumes of data.  Moreover, many Big Data frameworks have been recently developed such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
 \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives in high rate (velocity), and it comes in a mix of different data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.

\par The processing models of Big Data are Batch processing and Stream processing. Batch processing involves handling  large, finite volumes of data (e.g., processing of historical transaction records), and is more concerned with throughput. While Stream processing  is a method to manage fast and continuously incoming data (real-time), as it arrives to the system (e.g., credit card fraud detection and network monitoring applications). Stream processing systems usually  provide low latency processing. Both processing models are capable of operating with structured or semi-structured data (i.e., variety).

\par There exist some publications that compare Apache Spark and Apache Flink
 such as the Yahoo Streaming Benchmark \cite{yahoo}. But we believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
\item both frameworks are continuously improving in short time 
\item we aim to study the performance of some complex operations such as \texttt{groupByKey} \& \texttt{keyBy} in Apache Spark and Apache Flink, respectively.
\end{enumerate*}

\par The purpose of this report is to provide a comparative experimental evaluation of throughput and latency of stream processing in Apache Spark and Apache Flink. In order to simulate real-world use cases, we use datasets of airplane trajectories provided by the  Datacron project\footnote{\url{http://www.datacron-project.eu/}} to develop multiple streaming data processing workloads. Beside the performance comparison of the stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section~\ref{sec:sec3}) and how they can be solved in each platform. Informally,  this work aims to help the developers to determine which platform to use for different use cases.

\par The remainder of this report is organized as follows.
In Section~\ref{sec:sec2}, we present the main characteristics, programming APIs, and main data abstraction of Apache Spark and Apache Flink. Section~\ref{sec:sec3} presents the general aspects of stream processing tasks and the implementation of experimental workloads. Section~\ref{sec:sec4}
compares the performance results of the different workloads. And finally, Section~\ref{sec:sec5} gives the overall conclusion.

\section{Technical Background}
\label{sec:sec2}
 In this section, we present the architecture, key characteristics, programming APIs  of Apache Spark and Apache Flink. Moreover, a brief overview of the Apache Kafka platform is  provided.

@@

\par Apache Flink is an open source project that provides a large-scale, distributed stream processing platform \cite{flink}. Flink treats the batch processing as special case of streaming applications (i.e., finite stream).
It offers programming APIs in Java and Scala. Its software stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML)  and Flink Graph API (Gelly).

\par The main data abstraction of Flink are DataStream and DataSet that represent read-only collection of data elements. The list of elements is bounded (i.e., finite) in DataSet, while it is unbounded (i.e., infinite) in case of DataStream.

@@

\subsection{Data Stream Setup}
The evaluation streaming workloads read an input data stream from Kafka. In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that represent the aircraft's position over time. The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.  And  these datasets contain around \texttt{26 millions} messages.  
\par In our experiment a data stream producer component reads the ADS-B message datasets, then publishes them to the Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink. Figure~\ref{fig:data_stream} illustrates the data producer and the Kafka cluster setup in all experiments.
The data stream is divided into four portions over two server. The data producer publishes the stream records randomly to a Kafka partition and all ADS-B messages are published to the same topic (\texttt{datacorn}). Moreover, the data producer attaches the time of publishing to the ADS-B message line before publishing it to the Kafka topic.

@@

\newpage
\subsection{Evaluation Streaming Workloads}
In our experimental evaluation, we developed two real-time stream processing workloads that read a ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods. The workloads are designed to cover general key aspects of real-time stream processing tasks and evaluate the corresponding solutions in Spark and Flink. The following are the aspects of streaming data processing were covered by our workloads's design:

\begin{itemize}
\item Handling parallel input streams (e.g., Kafka Stream).
\item How to aggregate the state of an input stream.
\item Manage the order of stream records.
\item How to provide and update global data model in a stream processing task.
\item Evaluate the performance by measuring the latency and throughput.
\end{itemize}
We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Sections~\ref{sec:statistics} and~\ref{sec:sectors}. Afterward, In Section~\ref{sec:sec4} we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks. 
\par The two streaming workloads and the data stream producers are implemented in Java and the source code is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.

\subsubsection{Statistics Computation per Trajectory Workload}
\label{sec:statistics}
In the first stream processing workload, we construct a stream of trajectories by considering  the position messages (i.e., ADS-B messages) that belong to the same aircraft as trajectory. Moreover, we continuously compute and aggregate statistics for each new position in a trajectory . As example of the computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc. In this workload, we cover the parallel receiving of an input data stream, statefull aggregation over a data stream and preserving the correct order of the stream's records.
The implementation details of this workload in Spark and Flink are presented in the following: 
\begin{itemize}
\item {\bf{Implementation in Spark }}

\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream}, which represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, multiple instances of \texttt{DStream} must be created and combined together using a \texttt{union} operation. The output stream of the \texttt{union} is further processed using a \texttt{mapToPair} transformation that  maps each position message to tuple of aircraft ID and position message. Then,  irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation. Afterward, tuples that share the same aircraft ID are groped together to construct a stream of trajectories (tuples of ID and list of positions) by applying the \texttt{groupByKey} operation. Since the processing  model of Spark Streaming is micro-batch based and stateless an \texttt{updateStateByKey} operation must be used to maintain the state of the stream of trajectories. In the context of this workload, the state is the aggregated statistics of a trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that is continuously applied for every batch.
 The state update function uses the last position of a trajectory from previous batch to aggregate and compute statistics quantities to the received positions within a new batch and new positions are kept as the new state. Inside the state update function  the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.

@@

The details of Flink's implementation of statistics computation workload as follows. First, a \texttt{FlinkKafkaConsumer} is used to  construct a \texttt{DataStream} instance that represents the Kafka stream of aircrafts positions. The  \texttt{FlinkKafkaConsumer} implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka stream's partitions balanced between the parallel instances. Second, the stream's records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation. Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type. Fourth, the filtered tuples  are grouped based on the tuple's ID using a \texttt{keyBy} operator that constructs a \texttt{KeyedStream} of trajectories. Finally, 
a \texttt{reduce} transformation on the  \texttt{KeyedStream} is used to calculate the statistics for each new arriving trajectory's position by utilizing the aggregated statistics of the previous position. The old position is discarded and the new position with computed statistics is used to aggregate the statistics of a trajectory.

@@

\subsubsection{Air Sector Change Detection Workload}
\label{sec:sectors}
The goal of sector change detection workload is to detect the  entering or leaving of an aircraft from one air sector to another one, by processing the real-time stream of aircraft positions. Given the fact that a dataset of sectors (i.e., polygons) is available as a reference to assign the corresponding sector for certain aircraft's position. In this workload, we test the aspect of providing a global data model (sectors dataset) in stream processing workload. Moreover, the preserving of the stream's records order and consuming of parallel Kafka streams  are covered.

@@

 We implemented the sector change detection workload in Spark Streaming by constructing the stream of trajectories and assigning a sector for each position of a trajectory. A change between the assigned sectors of two consecutive aircraft's positions is marked  as a transition of sector for a certain trajectory. \par First, multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.create} \texttt{DirectStream}  are combined using \texttt{union} to read in  parallel from Kafka. Second, the records (i.e., ADS-B messages) of the resulting stream are parsed and mapped to tuples of (aircraft ID, position message) using \texttt{mapToPair} transformation. Third, the unrelated  tuples are filtered based on the message type using a \texttt{filter} operation. Fourth, the stream of trajectories is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function. Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied. The state update function assigns the corresponding sector for each new position, the old positions of a certain trajectory from the previous batch are discarded, but only the last position of the old batch is kept to be used alongside with the new positions. In addition to that the new positions are sorted by th streaming time to maintain the correct order.
 \par The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in an efficient matter. Finally, the detection of a sector change is done by an extra filtering transformation using a \texttt{filter} operator, which checks the sectors of consecutive positions in a trajectory to find the ones with a change in the sector assignment.

@@

 The Flink's implementation of sector change detection workload uses the \texttt{FlinkKafkaConsumer09} that reads the position messages (i.e., ADS-B) stream from Kafka, which is used to construct a \texttt{DataStream} instance. The \texttt{DataStream}'s records are mapped to tuples of (aircraft ID and aircraft's position) by a configured \texttt{map} transformation. Irrelevant tuples are filtered based on the message type using a \texttt{filter} operation. Afterward, the tuples that shares a common IDs are grouped together by the \texttt{keyBy} operator. Since the Flink processes the stream records as they come, a \texttt{reduce} operator assigns the corresponding sector of position on the new tuple, while the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple then the old tuple is discarded. Finally, the tuples (i.e., tuples of (aircraft ID, position message)) with different previous and current sectors are filtered using a \texttt{filter} transformation.

@@

 \par To sum up, in this workload we have the  same differences between Flink and Spark Streaming regarding the state management and parallel  consumers of Kafka stream as was discussed in Section~\ref{sec:statistics}. Another key difference between them is how  to provide a global data model (e.g., sectors) in a streaming workload. Spark Streaming offers the \texttt{Broadcast} feature to handle this issue in efficient way, while the global data model is manually provided  to the stream operations in Flink.
\section{Performance Results}
\label{sec:sec4}
\par This section presents the results of performance evaluation in Spark (2.0.2) and Flink (1.1.3) for the streaming workloads that were described in Section 3.2. Specifically, we provide  the measurements of latency and throughput of the benchmarking workloads execution in Spark Streaming and Flink. We carried the experiment by setting up the Kafka stream as described in Section 3.1 on a local machine\footnote{ The experiments have been carried  on single machine with Ubuntu 16.04 LTS os, Intel® Core™ i5-6200U CPU @ 2.30GHz × 4  processor, and 16 GB of memory.} and execute the different workloads on the same machine (i.e., local mode) to obtain the performance results.

\par We consider the time delay between the streaming time of a record in Kafka and  the time when the record is processed by a workload as the platform's processing latency (measured in milliseconds). And the number of processed stream's records (aircrafts position messages) per second within each streaming workload is the throughput measure. Since the Spark Streaming is micro-batch based,  we examine the effect of different batch intervals
of the Spark Streaming on the performance measurements.

@@

Figure~\ref{fig:latency} shows the results of the latency measurements of the streaming workloads in Spark and Flink. The presented values present the average value of multiple trails and measurements. Measurements for different batch durations are given in Spark Streaming, we noticed that making the batch time too small does not give better latency results. Results show that Flink outperform Spark streaming in term of processing latency in the two evaluation workloads.

@@

\label{sec:sec5}
In this report, we present an experimental evaluation of the stream processing performance and capabilities of Apache Spark and Apache Flink. We carried our experiment over aircraft positions datasets provided by the Datacorn project, we developed two streaming workloads to cover general aspects of real-time stream processing tasks. Additionally,  we measured the latency and throughput of those workloads solution of Spark and Flink. Results show that Flink outperforms Spark Streaming in term of processing latency. In contrast, Spark Streaming has showed better throughput rates than Flink. 
\par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and the batch interval comparing to Spark streaming. Moreover, Flink operates over the individual record of the data stream, while Spark Streaming  processes the input data stream as micro-batches. Thus we can state that  Flink is a true stream processing framework. A side remark, we have  noticed that finding online resources related to Spark is easier than to Flink.

In future work, we plan to carry out our experiments over cluster of computers not just only on single computer. In addition, we will cover more aspects of real-time data stream processing and developing more evaluation workloads.

>>>>

\end{abstract}
Recent years have witnessed advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For instance, both frameworks support real-time data 
stream processing. The question of which platform is superior to the other still open. 
In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink by measuring different performance metrics (latency and throughput). Furthermore, we cover some key aspects of streaming data applications and how they are handled in the two frameworks. The experiment is done using multiple streaming workloads over real-world datasets. Results show that Flink outperforms Spark streaming in term of latency. While Spark Streaming gives higher throughput rates than Flink. 
\section{Introduction}

\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia. This attention is a result of the fact that the amount of data we generate is increasing exponentially \cite{idc}.  Different sources such as  social media platforms, mobile phones, Internet of things (IoT), financial transactions, etc.  create large volumes of data.  Moreover, many Big Data frameworks have been recently developed such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
 \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives at a high rate (velocity), and it comes in a mix of different data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.

\par The processing models of Big Data are Batch processing and Stream processing. Batch processing involves handling  large and finite volumes of data (e.g., processing of historical transaction records), and is more concerned with throughput. While Stream processing  is a method to manage fast and continuously incoming data (real-time) as it arrives in the system (e.g., credit card fraud detection and network monitoring applications). Stream processing systems usually  provide low latency processing. Both processing models are capable of operating with structured or semi-structured data (i.e., variety).

\par There exist some publications that compare Apache Spark and Apache Flink
 such as the Yahoo Streaming Benchmark \cite{yahoo}. But we believe that we still need to carry out our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
\item both frameworks are continuously improving in a short time 
\item we aim to study the performance of some complex operations such as \texttt{groupByKey} \& \texttt{keyBy} in Apache Spark and Apache Flink, respectively.
\end{enumerate*}

\par The purpose of this report is to provide a comparative experimental evaluation of throughput and latency of stream processing in Apache Spark and Apache Flink. In order to simulate real-world use cases, we use datasets of airplane trajectories provided by the  Datacron project\footnote{\url{http://www.datacron-project.eu/}} to develop multiple streaming data processing workloads. Besides the performance comparison of the stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section~\ref{sec:sec3}) and how they can be solved in each platform. Informally,  this work aims to help the developers to determine which platform to use for different use cases.

\par The remainder of this report is organized as follows.
In Section~\ref{sec:sec2}, we present the main characteristics, programming APIs, and main data abstraction of Apache Spark and Apache Flink. Section~\ref{sec:sec3} presents the general aspects of stream processing tasks and the implementation of experimental workloads. Section~\ref{sec:sec4}
compares the performance results of the different workloads. And finally, Section~\ref{sec:sec5} gives the overall conclusion and future work.

\section{Technical Background}
\label{sec:sec2}
 In this section, we present the architecture, key characteristics and the programming APIs  of Apache Spark and Apache Flink. Moreover, a brief overview of the Apache Kafka platform is  provided.

@@

\par Apache Flink is an open source project that provides a large-scale, distributed stream processing platform \cite{flink}. Flink treats the batch processing as a special case of streaming applications (i.e., finite stream).
It offers programming APIs in Java and Scala. Its software stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML)  and Flink Graph API (Gelly).

\par The main data abstractions of Flink are DataStream and DataSet that represent read-only collection of data elements. The list of elements is bounded (i.e., finite) in DataSet, while it is unbounded (i.e., infinite) in the case of DataStream.

@@

\subsection{Data Stream Setup}
The evaluation streaming workloads read an input data stream from Kafka. In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance - Broadcast (ADS-B) messages that represent the aircraft's position over time. The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.  And  these datasets contain around \texttt{26 millions} messages.  
\par In our experiment a data stream producer component reads the ADS-B message datasets, then publishes them to the Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink. Figure~\ref{fig:data_stream} illustrates the data producer and the Kafka cluster setup in all experiments.
The data stream is divided into four portions over two servers. The data producer publishes the stream records randomly to a Kafka partition and all ADS-B messages are published to the same topic (\texttt{datacorn}). Moreover, the data producer attaches the time of publishing to the ADS-B message line before publishing it to the Kafka topic.

@@

\newpage
\subsection{Evaluation Streaming Workloads}
In our experimental evaluation, we developed two real-time stream processing workloads that read a stream of ADS-B messages from Kafka in order to perform basic trajectories analysis methods. The workloads are designed to cover general key aspects of real-time stream processing tasks and evaluate the corresponding solutions in Spark and Flink. The following are the aspects of streaming data processing were covered by our workloads design:

\begin{itemize}
\item Handling parallel input streams (e.g., Kafka Stream).
\item How to aggregate the state of an input stream.
\item Manage the order of stream records.
\item How to provide and update global data model in a stream processing task.
\item Evaluate the performance by measuring the latency and throughput.
\end{itemize}
We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Sections~\ref{sec:statistics} and~\ref{sec:sectors}. Afterward, In Section~\ref{sec:sec4} we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks. 
\par The two streaming workloads and the data stream producer are implemented in Java and the source code is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.

\subsubsection{Statistics Computation per Trajectory Workload}
\label{sec:statistics}
In the first stream processing workload, we construct a stream of trajectories by considering  the position messages (i.e., ADS-B messages) that belong to the same aircraft as trajectory. Moreover, we continuously compute and aggregate statistics for each new position in a trajectory. As an example of the computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc. In this workload, we cover the parallel receiving of an input data stream, stateful aggregation over a data stream and preserving the correct order of the stream's records.
The implementation details of this workload in Spark and Flink are presented in the following: 
\begin{itemize}
\item {\bf{Implementation in Spark }}

\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream}, which represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, multiple instances of \texttt{DStream} must be created and combined together using a \texttt{union} operation. The output stream of the \texttt{union} is further processed using a \texttt{mapToPair} transformation that  maps each position message to a tuple of aircraft ID and position message. Then,  irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation.\par Afterward, tuples that share the same aircraft ID are grouped together to construct a stream of trajectories (tuples of ID and list of positions) by applying the \texttt{groupByKey} operation. Since the processing  model of Spark Streaming is micro-batch based and stateless an \texttt{updateStateByKey} operation must be used to maintain the state of the stream of trajectories. In the context of this workload, the state is the aggregated statistics of a trajectory, so the statistics computation is performed within the custom \texttt{updateStateByKey} function that is continuously applied for every batch.
 \par The state update function uses the last position of a trajectory from previous batch to aggregate and compute statistics quantities to the received positions within a new batch and new positions are kept as the new state. Inside the state update function  the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.

@@

The details of Flink's implementation of statistics computation workload as follows. First, a \texttt{FlinkKafkaConsumer} is used to  construct a \texttt{DataStream} instance that represents the Kafka stream of aircrafts positions. The  \texttt{FlinkKafkaConsumer} implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka stream's partitions balanced between the parallel instances. Second, the stream's records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation. Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type. Fourth, the filtered tuples  are grouped based on the tuple's ID using a \texttt{keyBy} operator that constructs a \texttt{KeyedStream} of trajectories.\par Finally, 
a \texttt{reduce} transformation on the  \texttt{KeyedStream} is used to calculate the statistics for each new arriving trajectory's position by utilizing the aggregated statistics of the previous position. The old position is discarded and the new position with computed statistics is used to aggregate the statistics of a trajectory.

@@

\subsubsection{Air Sector Change Detection Workload}
\label{sec:sectors}
The goal of sector change detection workload is to detect the  entering or leaving of an aircraft from one air sector to another one, by processing the real-time stream of aircraft positions. Given the fact that a dataset of 20,000 sectors (i.e., polygons) is available as a reference to assign the corresponding sector for certain aircraft's position. In this workload, we test the aspect of providing a global data model (sectors dataset) in stream processing workload. Moreover, the preserving of the stream's records order and consuming of parallel Kafka streams  are covered.

@@

 We implemented the sector change detection workload in Spark Streaming by constructing the stream of trajectories and assigning a sector for each position of a trajectory. A change between the assigned sectors of two consecutive aircraft's positions is marked  as a sector's transition for a certain trajectory. \par First, multiple instances of \texttt{DStream} are created using the \texttt{KafkaUtils.create} \texttt{DirectStream} and combined using a \texttt{union} to read in  parallel from Kafka. Second, the records (i.e., ADS-B messages) of the resulting stream are parsed and mapped to tuples of (aircraft ID, position message) using \texttt{mapToPair} transformation. Third, the unrelated  tuples are filtered based on the message type using a \texttt{filter} operation. Fourth, the stream of trajectories is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function. Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied. The state update function assigns the corresponding sector for each new position, the old positions of a certain trajectory from the previous batch are discarded, but only the last position of the old batch is kept to be used alongside with the new positions. In addition to that the new positions are sorted by the streaming time to maintain the correct order.
 \par The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in an efficient matter. Finally, the detection of a sector change is done by an extra filtering transformation using a \texttt{filter} operator, which checks the sectors of consecutive positions in a trajectory to find the ones with a change in the sector assignment.

@@

 The Flink's implementation of sector change detection workload uses the \texttt{FlinkKafkaConsumer09} that reads the position messages (i.e., ADS-B) stream from Kafka, which is used to construct a \texttt{DataStream} instance. The \texttt{DataStream}'s records are mapped to tuples of (aircraft ID and aircraft's position) by a configured \texttt{map} transformation. Irrelevant tuples are filtered based on the message type using a \texttt{filter} operation. Afterward, the tuples that share a common ID are grouped together by the \texttt{keyBy} operator. Since the Flink processes the stream records as they come, a \texttt{reduce} operator assigns the corresponding sector of position on the new tuple, while the old tuple with the same ID is used to retrieve the previous sector to attach it to the new tuple then the old tuple is discarded. Finally, the tuples (i.e., tuples of (aircraft ID, position message)) with different previous and current sectors are filtered using a \texttt{filter} transformation.

@@

 \par To sum up, in this workload we have the  same differences between Flink and Spark Streaming regarding the state management and parallel  consumers of Kafka stream as was discussed in Section~\ref{sec:statistics}. Another key difference between them is how  to provide a global data model (e.g., sectors) in a streaming workload. Spark Streaming offers the \texttt{Broadcast} feature to handle this issue in an efficient way, while the global data model is manually provided  to the stream operations in Flink.
\section{Performance Results}
\label{sec:sec4}
\par This section presents the results of performance evaluation in Spark (2.0.2) and Flink (1.1.3) for the streaming workloads that were described in Section 3.2. Specifically, we provide  the measurements of latency and throughput of the benchmarking workloads execution in Spark Streaming and Flink. We carried the experiment by setting up the Kafka stream as described in Section 3.1 on a local machine\footnote{ The experiments have been carried  on single machine with Ubuntu 16.04 LTS os, Intel® Core™ i5-6200U CPU @ 2.30GHz × 4  processor, and 16 GB of memory.} and execute the different workloads on the same machine (i.e., local mode) to obtain the performance results.

\par We consider the time delay between the streaming time of a record in Kafka and  the time when the record is processed by a workload as the platform's processing latency (measured in milliseconds). And the number of processed stream's records (aircraft positions) per second within each streaming workload is the throughput measure. Since the Spark Streaming is micro-batch based,  we examine the effect of different batch intervals
of the Spark Streaming on the performance measurements.

@@

Figure~\ref{fig:latency} shows the results of the latency measurements of the streaming workloads in Spark and Flink. The presented values present the average value of multiple trails and measurements. Measurements for different batch durations are given in Spark Streaming, we noticed that making the batch time too small does not give better latency results. Results show that Flink outperforms Spark Streaming in term of processing latency in the two evaluation workloads.

@@

\label{sec:sec5}
In this report, we present an experimental evaluation of the stream processing performance and capabilities of Apache Spark and Apache Flink. We carried our experiment over aircraft positions datasets provided by the Datacorn project, we developed two streaming workloads to cover general aspects of real-time stream processing tasks. Additionally,  we measured the latency and throughput of those workloads solution of Spark and Flink. Results show that Flink outperforms Spark Streaming in term of processing latency. In contrast, Spark Streaming provides better throughput rates than Flink. 
\par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and the batch interval comparing to Spark streaming. Moreover, Flink operates over the individual record of the data stream, while Spark Streaming  processes the input data stream as micro-batches. Thus we can state that  Flink is a true stream processing framework. A side remark, we have  noticed that finding online resources related to Spark is easier than to Flink.

In future work, we plan to carry out our experiments over cluster of computers not just only on single computer. In addition, we will cover more aspects of real-time data stream processing and develop more evaluation workloads.
